|

### 2021-05-04 05-Chp-CPRD-Analysis.Rmd
End of day: -2856

@@

Cross reference here aswell re: the lack of information on Vascular dementia and the need for analyses at different sources of bias.

@@

<!--- Insert flow-chart here --->

@@

Using the Smeeth code-list, we get a similar breakdown of cases to that paper, with Alzheimer’s disease being a lot less frequent than when using our code-lists (which is not surprising based on the points above). <!-- This, in addition to unexplained statistical issues such as adjusting for post-baseline covariates, means I am sceptical of the value of comparing our findings with Smeeth. -->

@@

@@

@@

@@

@@

By pregnancy

By age-at-entry

By statin type (lipo vs hydrophilic)

@@

Backpain

Coronary artery disease

@@

@@

@@

In light of my own experiences in attempting to extract information for papers assessing prevantive measures, I attempted to document

This study was reported in line with the Strengthening the Reporting of Observational Studies in Epidemiology (STROBE) reporting guidelines[@vandenbroucke2007] (see Appendix \@ref() for the STROBE checklist) <!--- Maybe don't include the whole thing, or apply it to this thesis rather than the paper --->

Reporting guidelines do not always work, but help to guide

All code and summary statistics (i.e. the tables presented in this chapter, plus summary tables of effect estimates) are readily available in machine readable formats, without any formatting (i.e. as comma separated values, or CSV, files) from the Github repository for this project. <!--- Will need to cite this, and may need to archive on Zenodo. --->

@@

We have provided new evidence on the potential repurposing of lipid-regulating agents for the prevention of all-cause dementia, Alzheimer’s disease, vascular dementia, and other dementia. We found use of lipid-regulating agents not associated with probable or possible Alzheimer’s disease, but were associated with an increased risk of all-cause, vascular and other dementias. In all cases, the estimated associations were driven by those observed in the statin subgroup, which comprised the majority of participants in our cohort.

We have attempted to account for important sources of bias in our analysis and provide a comparison with other available literature. However, there is a strong potential for unmeasured confounding, misclassification and reverse causation, which raises questions about our findings, in particular the unexpected increase in risk of vascular dementia associated with statin use. Future research should aim to address these potential biases and, while it may be costly in terms of time and resources, a large scale, long-term randomized controlled trial would provide useful additional information on the effect of lipid-regulating agents on the risk of dementia and related outcomes.

>>>>

```{r load-files, echo = FALSE}
```

@@

Cross reference here as well re: the lack of information on Vascular dementia and the need for analyses at different sources of bias.

@@

@@

Using the Smeeth code-list, we get a similar breakdown of cases to that paper, with Alzheimer’s disease being a lot less frequent than when using our code-lists (which is not surprising based on the points above). <!-- This, in addition to unexplained statistical issues such as adjusting for post-baseline covariates, means I am sceptical of the value of comparing our findings with Smeeth. -->
IN addition, this analysis used propensity s. IN addition, an analytical approach which appeared to adjust for key variables likely to be on the analytical pathway between

@@

### Patient sample

<!----------------------------------------------------------------------------->
(ref:cprdFlowchart-cap) __Attrition of CPRD participants as the eligibility criteria were applied:__

(ref:cprdFlowchart-scap) Attrition of CPRD participants

```{r cprdFlowchart, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:cprdFlowchart-cap)', out.width='100%', fig.scap='(ref:cprdFlowchart-scap)'}
knitr::include_graphics(file.path("figures","cprd-analysis","cohort_attrition.png"))
```
<!----------------------------------------------------------------------------->

<!--- Add some text here about the fact highlighting the key criteria that excluded participants --->

@@

```{r characteristics, echo = FALSE}
```

Of the `r format(attrition[1,1], big.mark = ",")` participants included in our extract, `r format(attrition[14,1], big.mark = ",")` met the inclusion criteria (See Supplementary Figure 2 for the attrition flowchart), with a total follow-up of 11,091,437 patient years at risk. The median age at index was `r age_text` and participants were followed up for a median of `r fu_text`. During follow-up, an all-cause dementia diagnosis was recorded for `r format(p1[1,2], big.mark = ",")` patients  (`r format(p1[2,2], big.mark = ",")` probable AD, `r format(p1[3,2], big.mark = ",")` possible AD, `r format(p1[4,2], big.mark = ",")` vascular dementia, `r format(p1[5,2], big.mark = ",")` other dementia; see Supplementary Figure 1). The distribution of baseline characteristics across the drug classes can be seen in Table \@ref(tab:cprdCharacteristics-table).

<!--- Table 1 is missing type 2 diabetes --->

\blandscape
<!----------------------------------------------------------------------------->
(ref:cprdCharacteristics-caption) Patient characteristics by drug class

(ref:cprdCharacteristics-scaption) Patient characteristics by drug class

```{r cprdCharacteristics-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->
\elandscape

A substantial majority (`r percentage.statins`) of participants prescribed a lipid-regulating agent were prescribed a statin. We excluded the "Ezetimibe and statins" and "Nicotinic acid groups" classes from subsequent analysis based on the extremely small number of participants in these groups (Table \@ref(tab:cprdCharacteristics-table)). The "Ezetimibe and statins" treatment group represent those prescribed a treatment containing both ezetimibe and statins, rather than those where the two treats were prescribed concurrently.

@@

```{r missing-data, echo = FALSE}
```

```{r azd-text, echo = FALSE}
```

Full covariate information was available for `r missingtext`. Five key variables had some missing data: IMD 2010 score, a proxy for socioeconomic position that is measured as twentiles with 1 indicating the least deprived and 20 indicating the most deprived, was missing for `r imdtext`, because it is only recorded for English practices; alcohol status was missing for `r alcoholtext`; smoking status was missing for `r smokingtext`; BMI, or a calculated BMI from height and weight measurements, was missing for `r bmitext`, and baseline total cholesterol was missing for `r choltext`.

@@

All results described below are from the fully adjusted model. The results of the age-only model are presented in Supplementary Figure 3, and demonstrate that adjustment for additional covariates had a minimal effect.

__Alzheimer's disease__

As shown in Figure \@ref(fig:fig-results), our results show no effect of lipid-regulating agents on probable (`r probad_text`) and possible (`r possad_text`) Alzheimer's disease when compared with no treatment, with the exception of fibrates on probable Alzheimer's disease (`r probad_fib_text`).

__Non-Alzheimer's disease dementias__

In contrast to the findings for Alzheimer's disease outcomes, lipid-regulating agents were associated with an increased risk of a subsequent diagnosis of vascular dementia (`r vasdem_text`) or other dementia (`r othdem_text`). Again this effect was driven mainly by the statin subgroup, but there was some evidence that ezetimibe was associated with an increased risk of vascular (`r vasdem_eze_text`) and other (`r othdem_eze_text`) dementia.

__All-cause dementia__

For the composite all-cause dementia outcome, we found treatment with a lipid-regulating agent was associated with a slightly increased risk (`r anydem_text`), but the magnitude of the association was not as extreme as that observed for the vascular dementia subgroup. There was also some evidence that fibrates were associated with increased risk of all-cause dementia (`r anydem_fib_text`).

@@

When stratifying based on year of entry to the cohort, we observed no variation in risk by time period in any subgroup except for probable Alzheimer’s disease (Appendix XXX). Removing participants coded as female aged 55 and under at index from our analysis had minimal effect on our estimates (Appendix XXX).

<!-- For our negative control, there was some evidence that treatment with a lipid-regulating agent was associated with an increased risk of back pain (r sens_bp_text), suggesting there may be some residual confounding (Appendix XXX). -->

<!--- I actually think it would be better to include these results not in the appendix --->

<!-- By pregnancy -->

<!-- By age-at-entry -->

<!-- By statin type (lipo vs hydrophilic) -->

@@

<!-- Backpain -->

<!-- Coronary artery disease -->

_TBC - supervisors to email about it_

@@

<!--- To what extent should I cross-reference with the other literature here? --->

@@

Several previous studies have demonstrated that statins are the overwhelmingly favoured methods of cholesterol control. __[CITATION NEEDED]__

@@

In light of my own experiences in attempting to extract information for papers assessing preventative measures, as documented in Section \@ref(), the outputs from this analysis are described in detail in both. IN line with this, this analyses has been reported in line with the Strengthening the Reporting of Observational Studies in Epidemiology (STROBE) reporting guidelines[@vandenbroucke2007] (see Appendix \@ref() for the STROBE checklist) <!--- Maybe don't include the whole thing, or apply it to this thesis rather than the paper --->

Reporting guidelines do not always work, but help to guide accurate rep

All code and summary statistics (i.e. the tables presented in this chapter, plus summary tables of effect estimates) are readily available in machine readable formats (i.e. as comma separated values, or CSV, files) from the Github repository for this project. <!--- Will need to cite this, and may need to archive on Zenodo. --->

@@

This chapter has provided new evidence on the potential repurposing of lipid-regulating agents for the prevention of all-cause dementia, Alzheimer’s disease, vascular dementia, and other dementia. It found use of lipid-regulating agents associated with a small reduction in probable or possible Alzheimer’s disease, and were associated with an increased risk of all-cause, vascular and other dementias. In all cases, the estimated associations were driven by those observed in the statin subgroup, which comprised the majority of participants in our cohort.

This chapter attempted to account for important sources of bias and provide a comparison with other available literature, as identified in the systematic review presented in Chapter \@ref(sys-rev-heading). However, there is a strong potential for unmeasured confounding, misclassification and reverse causation, which raises questions about our findings, in particular the unexpected increase in risk of vascular dementia associated with statin use.

Given the observed increased association of statins in other

While these

Future research should aim to address these potential biases and, while it may be costly in terms of time and resources, a large scale, long-term randomized controlled trial would provide useful additional information on the effect of lipid-regulating agents on the risk of dementia and related outcomes.
